Bayesian Sparsification of Deep $\cplx$-valued networks

### 2020-03-18 paper/paper.tex
thorough proofread 2 (section 4) * added one-two column option to the docclass * improved the language a bit and compactified the narrative, removed duplicate comments * terminology: dropped log-uniform references, changed linear to dense linear * fixed eqns (5, 6), removed eqns (13', 16, 23, 24), swapped eqn 12 and 13 * verified ARD kl-div expressions

\documentclass[a4paper,10pt]{article}

@@

\textit{Stochastic Gradient Variational Bayes} (SGVB), proposed by \citet{kingma_auto-encoding_2014},
replaces the Variational objective \eqref{eq:elbo_general} by its unbiased differentiable
Monte-Carlo estimator, which makes it possible to employ stochastic gradient optimization
methods to solve \eqref{eq:variational-progam}. If $
  \omega \sim q_{\theta}
$ is equivalent in distribution to $
  \omega = g(\varepsilon; \theta)
$ for some differentiable $
  g(\varepsilon; \theta)
$ and non-parametric random variable $
    \varepsilon \sim p_\varepsilon
$, then the SGVB objective is
\begin{equation}  \label{eq:sgvb_objective}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta} \| \pi_{\lambda})
      % + \frac{N}{\lvert B \rvert} \sum_{i\in B}
        % \log p(z_i \mid \omega_i)b
      + \frac1{L} \sum_{l=1}^L
        \log p_{\phi}(D \mid g(\varepsilon_{l}; \theta))
        % \log p_{\phi}(D \mid \omega_{l})
        %   \Big\vert_{\omega_{l} = g(\varepsilon_{l}; \theta)}
      % =
      % - KL(q_{\theta} \| \pi)
      % + N \mathbb{E}_{B \sim \{D\}_M \otimes q_\theta^M}
        % \hat{\mathbb{E}}_{z,\omega \sim B}
          % \log p(z \mid \omega)
    \,, 
\end{equation}
where $
  (\varepsilon_{l})_{l=1}^L
$ is an iid sample from $p_\varepsilon$, the KL-divergence term replaced by unbiased
differentiable finite sample estimator if it is intractable.
%
For the purposes of this study, we assume that the likelihood satisfies $
  \log p_\phi(D\mid \omega)
    = \sum_i \log p_\phi(z_i \mid \omega)
$, i.e. the observed data $
  D = (z_i)_{i=1}^N
$ is conditionally independent, and consider \eqref{eq:sgvb_objective} with $L=1$
and random mini-batches from $D$:
\begin{equation}  \label{eq:elbo}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta} \| \pi_{\lambda})
      + \frac{N}{M} \sum_{k=1}^M
        \log p_{\phi}(z_{i_k} \mid g(\varepsilon_k; \theta))
    \,,
\end{equation}
for a random subsample $(z_{i_k})_{k=1}^M$ and iid $\varepsilon_k \sim p_\varepsilon$.
%
% (DSVI) titsias-lazaro.tex \todo{DSVI} \citep{titsias_doubly_2014}

@@

More recently general development of $\cplx$-valued networks has been continued by
\citet{trabelsi_deep_2017}, who outline the essential building blocks for deep $\cplx$-valued
networks and describe suitable representation and operations, including convolutional and
dense layers, $\cplx$-valued activations, complex batch-normalization and weight initialization.
Their approach coupled with $\cplx\real$ calculus for backprop transforms a $\cplx$-valued
network into an intricately connected $\real$-valued computational graph, constructured
to respects $\cplx$-arithmetic.
%
In particular, natural identification of $\cplx$ and $\real^2$ implies that linear layers $
  L \colon \cplx^m \to \cplx^n
$ act upon their inputs as follows:
\begin{equation}  \label{eq:cplx-lin-op}
  {(\real^n)}^{2}
    \to {(\real^m)}^{2}
    \colon (u, v)
      \mapsto \Bigl(
        P u - Q v,
        P v + Q u
      \Bigr)
    \,,
\end{equation}
where $
  P, Q \colon \real^{n} \to \real^{m}
$ are unique operators such that $L = P + j Q$. A $\cplx$-convolutional layer with kernel
$W$ can be implemented as two $\real$-convolutions with kernels $\Re{W}$ and $\Im{W}$ using
\eqref{eq:cplx-lin-op}. Activations include trigonometric functions and representation-dependent
non-linearities \citep{hirose_complex-valued_2009}: polar (phase-amplitude) $
  z \mapsto \sigma(\lvert z \rvert) e^{j \arg z}
$ or planar (real-imaginary) $
  z \mapsto \sigma(\Re z) + j \sigma(\Im z)
$ for some $\real$-valued non-linearity $\sigma$, possibly with learnable parameters
\citep{trabelsi_deep_2017,wolter_complex_2018}.

@@

In this section we review the complex Gaussian distribution and discuss its properties. Then
we introduce factorized dropout posterior for complex-valued parameters and the relevant
version of the local reparameterization trick. Then we move on to choosing suitable priors
and then onward to deriving explicit expressions for the divergence penalties in \eqref{eq:elbo}
in terms of special functions.

@@

%notation
In the following $M^{\top}$ is the matrix transpose, $\conj{M}$ is elementwise complex
conjugation, and $M^{\hop} = (\conj{M})^\top$ denotes Hermitian conjugate.

@@

$. Matrices $\Gamma$ and $C$ are $

@@

$, respectively.

@@

% $\cplx$-Gaussians are closed under affine transformations: if $y = A z + b$
% for $A \in \cplx^{n \times m}$ and $b \in \cplx^{n}$, then $
%   y \sim \mathcal{N}_n^{\cplx}\bigl(
%       A\mu + b, A \Gamma A^{\hop}, A C A^\top
%     \bigr)
% $.
$\cplx$-Gaussian random variables are closed under affine transformations: for $
  A \in \cplx^{n \times m}
$ and $b \in \cplx^{n}$
\begin{equation}  \label{eq:cn-affine}
  A z + b \sim \mathcal{N}_n^{\cplx}\bigl(
      A\mu + b, A \Gamma A^{\hop}, A C A^\top
    \bigr)
The entropy of a $\cplx$-Gaussian in terms of $\Gamma$ and $C$ is
\begin{align}  \label{eq:c-gauss-entropy-derivation}

@@

@@

@@

A $\cplx$-Gaussian vector $z$ is \textit{proper} if $z$ and $\conj{z}$ are uncorrelated, i.e.
$C = 0$. From \eqref{eq:c-gauss-entropy-derivation} its entropy is
\begin{equation}  \label{eq:cn-proper-entropy}
    = \tfrac12 \log \det{(\pi e \Gamma)} \det{(\pi e \conj{\Gamma})}
    \tag{\ref{eq:c-gauss-entropy-derivation}'}

@@

The local reparameterization trick provides both faster computations, and better statistical
efficiency of the SGVB gradient estimator \citep{wang_fast_2013,kingma_variational_2015}.
Suppose that weights $W \in \cplx^{n\times m}$ of a linear layer are drawn from a mean
field $\cplx$-Gaussian approximation
\begin{equation}  \label{eq:c-gauss-vi-general}
  W_{ij}
    \sim \mathcal{N}^{\cplx} \bigl(
      \mu_{ij}, \Sigma_{ij}, C_{ij}
    \bigr)
  \,,
\end{equation}
where mean, \textit{variance} and \textit{relation} are given by matrices $
  \mu, \Sigma, C \in \cplx^{n\times m}
$ with $\Sigma_{ij} \in \real$, and $
  \lvert C_{ij} \rvert^2 \leq \Sigma_{ij}
$. The property \eqref{eq:cn-affine} and independence of elements of $W$ imply that for any
$x \in \cplx^m$ and $b \in \cplx^n$ the components of $y = W x + b$ are independent with
distributions
\begin{equation}  \label{eq:cplx-gauss-trick}
  y_i
    \sim \mathcal{N}^{\cplx}
      \Bigl(
        % e_i^\top \mu x + b_i,
        b_i + \sum_{j=1}^m \mu_{ij} x_j,
        \, \sum_{j=1}^m \Sigma_{ij} \lvert x_j \rvert^2,
        \, \sum_{j=1}^m C_{ij} x_j^2
      \Bigr)
    \,.
\end{equation}
The trick requires three matrix operations: $\mu x + b$, $\Sigma \lvert x \rvert^2$ and
$C x^2$ with the modulus and power applied elementwise.

The trick and dropout can be applied to any layer, the output of which depends linearly on
its parameters, e.g. convolutional layers, affine, and bilinear transformations $
  (x, z) \mapsto x^\top W^{(j)} z + b_j
$. Similar to $\real$ case, variational dropout for $\cplx$ convolutions draws independent
realizations of $W$ for each spatial patch in the input. The rationale is simpler computations
and variance reduction. The latter effect stems from decorrelated gradient estimates from
overlapping patches \citep{kingma_variational_2015}, which allows \eqref{eq:cplx-gauss-trick}
to use $\cplx$ convolutions of elementwise complex squares $x^2$ and amplitudes $\lvert x \rvert^2$
with the relation and variance kernels.
% Weight sharing is essentially pushed to shared variational parameters $\theta$ in the approximation $q_\theta$.
% A convolution is a matrix product of $W$, unrolled into a toeplitz matrix, and im-to-col
% flattened intput $x$

@@

For $\cplx$-variational dropout we propose using mean-field proper $\cplx$-Gaussian approximate
posterior and MAP estimates of bias terms, i.e. parameters $W \in \cplx^{n\times m}$ of a dense
linear layer follow
% We use MAP estimates of bias terms, i.e. assume an independent fully factorized Dirac approximation.
\begin{equation}  \label{eq:c-gauss-vi}
  % (we omit the subscript $m=1$ for brevity)
  W \sim q(W)
    % = \prod_{ij} q(W_{ij})
    = \prod_{ij} \mathcal{N}^{\cplx}(
      W_{ij} \mid
        \mu_{ij},
        \alpha_{ij} \lvert \mu_{ij} \rvert^2,
        % \sigma^2_{ij},
        0
    )
  \,,
\end{equation}
and we use similar factorized approximation for convolutions and other, effectively, linear
layers. For this approximation the relation coefficients \eqref{eq:cplx-gauss-trick} are
zero.

We consider two priors for Bayesian dropout: \verify{log-uniform} and $\cplx$-Gaussian empirical
Bayes \citep{kingma_variational_2015,molchanov_variational_2017,kharitonov_variational_2018}.
For a fully factorized approximation $q(\omega)$ and factorized prior belief $\pi(\omega)$,
the divergence term in ELBO \eqref{eq:elbo} is
\begin{equation}  \label{eq:elbo-general-kl-div}
  KL(q \| \pi)
    % = \mathbb{E}_{\oemga \sim q(\oemga)} \log \tfrac{q(\oemga)}{\pi(\oemga)}
    % = \sum_{ij} \mathbb{E}_{\omega_{ij} \sim q(\omega_{ij})}
    %   \log \tfrac{q(\omega_{ij})}{\pi(\omega_{ij})}
    % = \sum_{ij} KL\bigl( q(\omega_{ij}) \| \pi(\omega_{ij}) \bigr)
    = - \sum_{ij}
        \mathbb{H}(q(\omega_{ij}))
        + \mathbb{E}_{q(\omega_{ij})} \log{\pi(\omega_{ij})}
    \,.
\end{equation}
We omit subscripts ${ij}$ for brevity in the next sections.

\subsubsection{Variational dropout} % (fold)
\label{ssub:variational_dropout}

From \eqref{eq:cn-proper-entropy} the KL-divergence for \verify{log-uniform prior} $
  \pi(\omega) \propto {\lvert \omega \rvert}^{-\beta}
$ with $\beta \geq 1$ is
\begin{equation}  \label{eq:c-vd-kl-div-raw}
  KL(q\| \pi)
    \propto
    %   \mathbb{E}_{\omega \sim q(\omega)} \log q(\omega)
    %   + \tfrac{\beta}2 \mathbb{E}_{\omega \sim q(\omega)} \log \lvert \omega \rvert^2
    % =
    %   - \log \lvert \det{(\pi e \sigma^2)} \rvert
    %   + \tfrac{\beta}2 \mathbb{E}_{\omega \sim q(\omega)} \log \lvert \omega \rvert^2
    % =
      - \log{\sigma^2}
      + \tfrac{\beta}2 \mathbb{E}_{\omega \sim q(\omega)} \log \lvert \omega \rvert^2
Property \eqref{eq:cn-affine} implies that $
  \mathcal{N}^{\cplx}(\mu, \sigma^2, 0)
  \sim \mu \cdot \mathcal{N}^{\cplx}(1, \alpha, 0)
$ for $\mu \neq 0$ and $
  \sigma^2 = \alpha \lvert \mu \rvert^2
$, whence
\begin{equation}  \label{eq:expect-improper-term-cplx}
  \mathbb{E}_{\omega \sim q(\omega)} \log \lvert \omega \rvert^2
    % = \mathbb{E}_{\xi \sim \mathcal{N}^{\cplx}(\xi \vert 1, \alpha, 0)}
    %   \log \lvert \mu \xi \rvert^2
    % = \log \lvert \mu \rvert^2
    %   + \mathbb{E}_{\xi \sim \mathcal{N}^{\cplx}(\xi \vert 1, \alpha, 0)}
    %       \log \lvert \xi \rvert^2
    = \log \alpha \lvert \mu \rvert^2
      + \mathbb{E}_{\xi \sim \mathcal{N}^{\cplx}(\xi \vert 0, 1, 0)}
          \log{\bigl\lvert \tfrac1{\sqrt{\alpha}} + \xi \bigr\rvert^2}
Let $
  (z_i)_{i=1}^m \sim \mathcal{N}^{\cplx}(0, 1, 0)
$ iid and $\theta \in \cplx^m$. Then $
  \sum_i \lvert \theta_i + z_i \rvert^2
    \sim \chi^2_{2m}(s^2)
$ with $
  s^2 = \sum_i \lvert \theta_i \rvert^2
$, i.e. a non-central $\chi^2_{2m}$ with parameter $s^2$. Its log-moments for general
integer $m \geq1$ are given in \cite[p.~2466]{lapidoth_capacity_2003}. In particular,
% in fact Appendix X, Lemma 10.1
for $m=1$ and $\theta\in \cplx$ we have
\begin{equation}  \label{eq:log-moment-for-chi-2}
  \mathbb{E}_{z \sim \mathcal{N}^{\cplx}(0, 1, 0)}
    \log \lvert \theta + z \rvert^2
    = \log \lvert \theta \rvert^2 - \mathop{Ei}( - \lvert \theta \rvert^2)
where $
  \mathop{Ei}(x) = \int^x_{-\infty} t^{-1} e^t dt
$ for $x < 0$ is the Exponential Integral. Note that
$
  \tfrac{d}{dx} \mathop{Ei}(x) = \tfrac{e^x}{x}
  \mathop{Ei}(x) \leq \log{(-x)}
  % as x to 0 from below
  \mathop{Ei}(x) \approx \log{(-x)} - \gamma
$ as $x\to 0$ ($\gamma$ is Euler's constant) and $
  \mathop{Ei}(x) \geq -e^x
$ for $x \leq -1$.
% 1. by Leibniz rule $Ei(x)$ has -ve derivative on $x < 0$
% 2. $Ei(x) \leq 0$ on $x < 0$, since $t^{-1} e^t \leq t^{-1} < 0$ on $t < 0$
% 3. $Ei(x) \leq \log{(-x)}$ on $x < 0$, since on $x \in [-1, 0)$
%   \mathop{Ei}(x)
%     = \mathop{Ei}(-1) + \int_{-1}^x t^{-1} e^t dt
%     \leq \mathop{Ei}(-1) + \int_{-1}^x t^{-1} dt
%     = \mathop{Ei}(-1) + \log{(-x)}
%     \leq \log{(-x)}
% 3. by l'H{\^o}pital rule $Ei(x)$ is asymptotically $\log{-x}$ as $x \to 0-$
% % t = -e^{-x} ,  -\log{(-t)} = x
%   \lim_{x\to 0-} \frac{\mathop{Ei}(x)}{\log{(-x)}}
%     = \lim_{x\to 0-} \frac{x^{-1} e^x}{x^{-1}}
%     = 1
% 4. \citep{lapidoth_capacity_2003} yields $\lim_{x\to 0-} \log{(-x)} - \mathop{Ei}(x) = \gamma$
% 5. $Ei(x) \geq -e^x$ for any $x < -1$, since $t^{-1} e^t \geq - e^t$ for $t < -1$
% \mathop{Ei}(x)
%   = \int_{-\infty}^x t^{-1} e^t dt
%   \geq \int_{-\infty}^x -e^t dt = -e^x
% 6. $Ei(x) \leq \log{(-x)} - \gamma$ on $x < 0$ and $-e^x \leq Ei(x) \leq 0$ on $x < -1$.
Through \eqref{eq:expect-improper-term-cplx} and \eqref{eq:log-moment-for-chi-2} we get
\begin{equation}  \label{eq:c-vd-kl-div}
  KL(q\| \pi)
    \propto
    %   - \log \lvert \det{(\pi e \sigma^2)} \rvert
    %   + \tfrac{\beta}2 \log \alpha \lvert \mu \rvert^2
    %   + \tfrac{\beta}2 \mathbb{E}_{\xi \sim \mathcal{N}^{\cplx}(\xi \vert 0, 1, 0)}
    %       \log{\bigl\lvert \tfrac1{\sqrt{\alpha}} + \xi \bigr\rvert^2}
    % =
    %   - \log{\sigma^2}  % - \log{\pi e}
    %   + \tfrac{\beta}2 \log \alpha \lvert \mu \rvert^2
    %   + \tfrac{\beta}2 \bigl(
    %     - \log \alpha - \mathop{Ei}( -\tfrac1{\alpha})
    %   \bigr)
    % =
    %   - \log{\sigma^2}
    %   + \tfrac{\beta}2 \log \alpha \lvert \mu \rvert^2
    %   - \tfrac{\beta}2 \log \alpha
    %   - \tfrac{\beta}2 \mathop{Ei}( -\tfrac1{\alpha})
    % =
      % \log \tfrac{\lvert \mu \rvert^\beta}{\sigma^2}
      % \log \tfrac{\lvert \mu \rvert^\beta}{\alpha \lvert \mu \rvert^2}
      \tfrac{\beta-2}2 \log{\lvert \mu \rvert^2}
      + \log{\tfrac1{\alpha}}
      - \tfrac{\beta}2 \mathop{Ei}(- \tfrac1{\alpha})
      \tag{\ref{eq:c-vd-kl-div-raw}'} \,.
We set $\beta = 2$ to make the divergence term depend only on $\alpha$ and add $\gamma$

>>>>

\documentclass[a4paper,10pt,onecolumn]{article}

@@

\textit{Stochastic Gradient Variational Bayes} (SGVB), proposed by \citet{kingma_auto-encoding_2014},
replaces the Variational objective \eqref{eq:elbo_general} by its unbiased differentiable
Monte-Carlo estimator, which makes it possible to employ stochastic gradient optimization
methods to solve \eqref{eq:variational-progam}. If $
  \omega \sim q_{\theta}
$ is equivalent in distribution to $
  \omega = g(\varepsilon; \theta)
$ for some differentiable $
  g(\varepsilon; \theta)
$ and non-parametric random variable $
    \varepsilon \sim p_\varepsilon
$, then the SGVB objective is
\begin{equation}  \label{eq:sgvb_objective}
  \widetilde{\mathcal{L}} %(\theta; \phi, \lambda)
    = - KL(q_{\theta} \| \pi_{\lambda})
      % + \frac{N}{\lvert B \rvert} \sum_{i\in B}
        % \log p(z_i \mid \omega_i)b
      + \frac1{L} \sum_{l=1}^L
        \log p_{\phi}(D \mid g(\varepsilon_{l}; \theta))
        % \log p_{\phi}(D \mid \omega_{l})
        %   \Big\vert_{\omega_{l} = g(\varepsilon_{l}; \theta)}
      % =
      % - KL(q_{\theta} \| \pi)
      % + N \mathbb{E}_{B \sim \{D\}_M \otimes q_\theta^M}
        % \hat{\mathbb{E}}_{z,\omega \sim B}
          % \log p(z \mid \omega)
    \,, 
\end{equation}
where $
  (\varepsilon_{l})_{l=1}^L
$ is an iid sample from $p_\varepsilon$, the KL-divergence term replaced by unbiased
differentiable finite sample estimator if it is intractable.
%
For the purposes of this study, we assume that the likelihood satisfies $
  \log p_\phi(D\mid \omega)
    = \sum_i \log p_\phi(z_i \mid \omega)
$, i.e. the observed data $
  D = (z_i)_{i=1}^N
$ is conditionally independent, and consider \eqref{eq:sgvb_objective} with $L=1$
and random mini-batches from $D$:
\begin{equation}  \label{eq:elbo}
  \widetilde{\mathcal{L}} %(\theta; \phi, \lambda)
    = - KL(q_{\theta} \| \pi_{\lambda})
      + \frac{N}{M} \sum_{k=1}^M
        \log p_{\phi}(z_{i_k} \mid g(\varepsilon_k; \theta))
    \,,
\end{equation}
for a random subsample $(z_{i_k})_{k=1}^M$ and iid $\varepsilon_k \sim p_\varepsilon$.
%
% (DSVI) titsias-lazaro.tex \todo{DSVI} \citep{titsias_doubly_2014}

@@

More recently general development of $\cplx$-valued networks has been continued by
\citet{trabelsi_deep_2017}, who outline the essential building blocks for deep $\cplx$-valued
networks and describe suitable representation and operations, including convolutional and
dense layers, $\cplx$-valued activations, complex batch-normalization and weight initialization.
Their approach coupled with $\cplx\real$ calculus for backprop transforms a $\cplx$-valued
network into an intricately connected $\real$-valued computational graph, constructured
to respects $\cplx$-arithmetic.
%
In particular, natural identification of $\cplx$ and $\real^2$ implies that dense linear
layers $
  L \colon \cplx^m \to \cplx^n
$ act upon their inputs as follows:
\begin{equation}  \label{eq:cplx-lin-op}
  {(\real^n)}^{2}
    \to {(\real^m)}^{2}
    \colon (u, v)
      \mapsto \Bigl(
        P u - Q v,
        P v + Q u
      \Bigr)
    \,,
\end{equation}
where $
  P, Q \colon \real^{n} \to \real^{m}
$ are unique operators such that $L = P + j Q$. A $\cplx$-convolutional layer with kernel
$W$ can be implemented as two $\real$-convolutions with kernels $\Re{W}$ and $\Im{W}$ using
\eqref{eq:cplx-lin-op}. Activations include trigonometric functions and representation-dependent
non-linearities \citep{hirose_complex-valued_2009}: polar (phase-amplitude) $
  z \mapsto \sigma(\lvert z \rvert) e^{j \arg z}
$ or planar (real-imaginary) $
  z \mapsto \sigma(\Re z) + j \sigma(\Im z)
$ for some $\real$-valued non-linearity $\sigma$, possibly with learnable parameters
\citep{trabelsi_deep_2017,wolter_complex_2018}.

@@

In this section we review the complex Gaussian distribution and discuss its properties. Then
we introduce factorized dropout posterior for complex-valued parameters and the relevant
version of the local reparameterization trick. Then we move on to choosing suitable priors
and then to deriving explicit expressions for the divergence penalties in \eqref{eq:elbo}
in terms of \verify{special} functions.

@@

@@

% notation
Here, $M^{\top}$ is the matrix transpose, $\conj{M}$ is elementwise complex
conjugation, and $M^{\hop} = (\conj{M})^\top$ denotes Hermitian conjugate.
Matrices $\Gamma$ and $C$ are given by $

@@

$, respectively, and $z$ is a \textit{proper} $\cplx$-Gaussian vector if $z$ and
$\conj{z}$ are uncorrelated, i.e. $C = 0$.

@@

The entropy of $z$ terms of $\Gamma$ and $C$ is
\begin{align}

@@

  % \label{eq:c-gauss-entropy-derivation}

@@

  \label{eq:cn-proper-entropy}
    % = \tfrac12 \log \det{(\pi e \Gamma)} \det{(\pi e \conj{\Gamma})}
    \Bigl[
    \quad \text{for } C = 0
    \Bigr]

@@

@@

The $\cplx$-variant of the local reparameterization trick utilizes the fact that Gaussian
vectors are closed under affine transformations:
% $\cplx$-Gaussians are closed under affine transformations
\begin{equation}  \label{eq:cn-affine}
  b + A z \sim \mathcal{N}_n^{\cplx}\bigl(
      b + A\mu, A \Gamma A^{\hop}, A C A^\top
    \bigr)
  \,,
\end{equation}
for $A \in \cplx^{n \times m}$ and $b \in \cplx^{n}$. Therefore, if $W \in \cplx^{n\times m}$
is drawn from a fully factorized approximation
\begin{equation}  \label{eq:c-gauss-vi-general}
  W \sim q(W)
    = \prod_{ij} \mathcal{N}^{\cplx} \bigl(
      \mu_{ij}, \Sigma_{ij}, C_{ij}
    \bigr)
  \,,
\end{equation}
with $
  \mu, C \in \cplx^{n\times m}
$, $\Sigma \in \real^{n\times m}$, and $
  \lvert C_{ij} \rvert^2 \leq \Sigma_{ij}
$, then for $x \in \cplx^m$ and $b \in \cplx^n$ the vector $y = b + W x$ has independent
components with
\begin{equation}  \label{eq:cplx-gauss-trick}
  y_i
    \sim \mathcal{N}^{\cplx}
      \Bigl(
        % e_i^\top \mu x + b_i,
        b_i + \sum_{j=1}^m \mu_{ij} x_j,
        \, \sum_{j=1}^m \Sigma_{ij} \lvert x_j \rvert^2,
        \, \sum_{j=1}^m C_{ij} x_j^2
      \Bigr)
    \,.
\end{equation}
The trick requires three operations, i.e. $b + \mu x$, $\Sigma \lvert x \rvert^2$ and
$C x^2$ with the complex modulus and square applied elementwise, and can be applied to
any layer, the output of which depends linearly on its parameters, such as convolutional,
affine, and bilinear transformations ($
  (x, z) \mapsto x^\top W^{(j)} z + b_j
$). Similar to $\real$ case, Variational Dropout for $\cplx$ convolutions draws independent
realizations of $W$ for each spatial patch in the input \citep{molchanov_variational_2017}.
This provides faster computations and better statistical efficiency of the SGVB gradient
estimator stemming from decorrelated gradient estimates from overlapping patches
\citep{kingma_variational_2015}, allowing \eqref{eq:cplx-gauss-trick} to leverage $\cplx$
convolutions of the relation and variance kernels with elementwise complex squares $x^2$
and amplitudes $\lvert x \rvert^2$.
% Weight sharing is essentially pushed to shared variational parameters $\theta$ in the approximation $q_\theta$.
% A convolution is a matrix product of $W$, unrolled into a toeplitz matrix, and im-to-col
% flattened intput $x$

@@

For $\cplx$-Bayesian Dropout we propose to use fully factorized proper $\cplx$-Gaussian
approximation \eqref{eq:c-gauss-vi-general} for weights and point estimates of biases in
dense linear, convolutional and other, effectively, affine layers, i.e. we let $
  \Sigma_{ij} = \alpha_{ij} \lvert \mu_{ij} \rvert^2
$ and fix $C_{ij} = 0$. Under this approximation the relation coefficients in \eqref{eq:cplx-gauss-trick}
are zero.
%
For a fully factorized approximation $q(\omega)$ and factorized prior belief $
  \pi(\omega) = \prod_{ij} \pi(\omega_{ij})
$,
the divergence term in ELBO \eqref{eq:elbo} is
\begin{equation}  \label{eq:elbo-general-kl-div}
  KL(q \| \pi)
    % = \mathbb{E}_{\oemga \sim q(\oemga)} \log \tfrac{q(\oemga)}{\pi(\oemga)}
    % = \sum_{ij} \mathbb{E}_{\omega_{ij} \sim q(\omega_{ij})}
    %   \log \tfrac{q(\omega_{ij})}{\pi(\omega_{ij})}
    % = \sum_{ij} KL\bigl( q(\omega_{ij}) \| \pi(\omega_{ij}) \bigr)
    = - \sum_{ij}
        \mathbb{H}(q(\omega_{ij}))
        + \mathbb{E}_{q(\omega_{ij})} \log{\pi(\omega_{ij})}
    \,.
\end{equation}
We consider two fully factorized priors: an improper prior for Variational Dropout and
$\cplx$-Gaussian ARD prior. We omit subscripts ${ij}$ for brevity in the next sections.

\subsubsection{VD prior} % (fold)
\label{ssub:vd_prior}

From \eqref{eq:cn-proper-entropy} the KL-divergence for an improper prior $
  \pi(\omega) \propto {\lvert \omega \rvert}^{-\beta}
$ with $\beta \geq 1$ is
\begin{equation}  \label{eq:c-vd-kl-div-raw}
  KL(q\| \pi)
    \propto
    %   \mathbb{E}_{\omega \sim q(\omega)} \log q(\omega)
    %   + \tfrac{\beta}2 \mathbb{E}_{\omega \sim q(\omega)} \log \lvert \omega \rvert^2
    % =
    %   - \log \lvert \det{(\pi e \sigma^2)} \rvert
    %   + \tfrac{\beta}2 \mathbb{E}_{\omega \sim q(\omega)} \log \lvert \omega \rvert^2
    % =
      - \log{\sigma^2}
      + \tfrac{\beta}2 \Bigl(
        \mathbb{E}_{\omega \sim q(\omega)} \log \lvert \omega \rvert^2
      \Bigr)
    \,.
\end{equation}
Property \eqref{eq:cn-affine} implies that $
  \mathcal{N}^{\cplx}(\mu, \sigma^2, 0)
  \sim \mu \cdot \mathcal{N}^{\cplx}(1, \alpha, 0)
$ for $\mu \neq 0$ and $
  \sigma^2 = \alpha \lvert \mu \rvert^2
$, whence the expectation in brackets is given by
\begin{equation}  \label{eq:expect-improper-term-cplx}
  % \mathbb{E}_{\omega \sim q(\omega)} \log \lvert \omega \rvert^2
  %   % = \mathbb{E}_{\xi \sim \mathcal{N}^{\cplx}(\xi \vert 1, \alpha, 0)}
  %   %   \log \lvert \mu \xi \rvert^2
  %   % = \log \lvert \mu \rvert^2
  %   %   + \mathbb{E}_{\xi \sim \mathcal{N}^{\cplx}(\xi \vert 1, \alpha, 0)}
  %   %       \log \lvert \xi \rvert^2
  %   =
    \log \alpha \lvert \mu \rvert^2
      + \mathbb{E}_{\xi \sim \mathcal{N}^{\cplx}(\xi \vert 0, 1, 0)}
          \log{\bigl\lvert \tfrac1{\sqrt{\alpha}} + \xi \bigr\rvert^2}
    \,.
\end{equation}
If $
  (z_i)_{i=1}^m
    \sim \mathcal{N}^{\cplx}(0, 1, 0)
$ iid and $\theta \in \cplx^m$, then $
  \sum_i \lvert \theta_i + z_i \rvert^2
    \sim \chi^2_{2m}(s^2)
$ with $
  s^2 = \sum_i \lvert \theta_i \rvert^2
$, i.e. a non-central $\chi^2_{2m}$ with parameter $s^2$. Its log-moments for general
integer $m \geq1$ have been derived by \citet[p.~2466]{lapidoth_capacity_2003}.
% in fact Appendix X, Lemma 10.1
In particular, for $m=1$ and $\theta\in \cplx$ we have
\begin{equation}  \label{eq:log-moment-for-chi-2}
  \mathbb{E}_{z \sim \mathcal{N}^{\cplx}(0, 1, 0)}
    \log \lvert \theta + z \rvert^2
    = \log \lvert \theta \rvert^2 - \mathop{Ei}( - \lvert \theta \rvert^2)
    \,,
\end{equation}
where $
  \mathop{Ei}(x) = \int^x_{-\infty} t^{-1} e^t dt
$ for $x < 0$ is the Exponential Integral, which satisfies $
  \mathop{Ei}(x) \leq \log{(-x)}
$, $
  % as x to 0 from below
  \mathop{Ei}(x) \approx \log{(-x)} - \gamma
$ as $x\to 0$ ($\gamma$ is Euler's constant) and $
  \mathop{Ei}(x) \geq -e^x
$ for $x \leq -1$.
% 1. by Leibniz rule $Ei(x)$ has -ve derivative on $x < 0$
% 2. $Ei(x) \leq 0$ on $x < 0$, since $t^{-1} e^t \leq t^{-1} < 0$ on $t < 0$
% 3. $Ei(x) \leq \log{(-x)}$ on $x < 0$, since on $x \in [-1, 0)$
% $$
%   \mathop{Ei}(x)
%     = \mathop{Ei}(-1) + \int_{-1}^x t^{-1} e^t dt
%     \leq \mathop{Ei}(-1) + \int_{-1}^x t^{-1} dt
%     = \mathop{Ei}(-1) + \log{(-x)}
%     \leq \log{(-x)}
%   \,. $$
% 3. by l'H{\^o}pital rule $Ei(x)$ is asymptotically $\log{-x}$ as $x \to 0-$
% $$
% % t = -e^{-x} ,  -\log{(-t)} = x
%   \lim_{x\to 0-} \frac{\mathop{Ei}(x)}{\log{(-x)}}
%     = \lim_{x\to 0-} \frac{x^{-1} e^x}{x^{-1}}
%     = 1
%   \,. $$
% 4. \citep{lapidoth_capacity_2003} yields $\lim_{x\to 0-} \log{(-x)} - \mathop{Ei}(x) = \gamma$
% 5. $Ei(x) \geq -e^x$ for any $x < -1$, since $t^{-1} e^t \geq - e^t$ for $t < -1$
% $$
% \mathop{Ei}(x)
%   = \int_{-\infty}^x t^{-1} e^t dt
%   \geq \int_{-\infty}^x -e^t dt = -e^x
%   \,. $$
% 6. $Ei(x) \leq \log{(-x)} - \gamma$ on $x < 0$ and $-e^x \leq Ei(x) \leq 0$ on $x < -1$.
%
Although $\mathop{Ei}$ is an intractable integral, requiring numerical approximations to
compute, its derivative is exact: $
  \tfrac{d}{dx} \mathop{Ei}(x) = \tfrac{e^x}{x}
$ at $x < 0$.
